---
title: "Georgia Pacific Case Study"
author: "Jensen brock"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup-libraries-data, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

# Load necessary Libraries
library(tidyverse)
library(broom)
library(scales)
library(modelr)
library(tidyquant)
library(forcats) 
library(gghalves)
library(lubridate)
library(scales)
library(car)
library(psych)

# Load Datasets
dataset_1 <- read_csv("data/dataset_1")
dataset_2 <- read_csv("data/dataset_2")
```


```{r data-cleaning, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# change column names to something better suited to coding
data1 <- dataset_1 %>% 
        rename(date = Date
             , price = `SRP ($)`
             , volume = `Volume (Units)`
             , revenue = `Sales ($)`
             )

# change column names to something better suited to coding
data2 <- dataset_2 %>% 
  rename(date = Date
       , price = `SRP ($)`
       , volume = `Volume (Units)`
       , revenue = `Sales ($)`
       , num_stores = `No of Stores`
       , shelf_space_pct = `Additional Shelf Space (%)`
       , is_advertisement = `Any Store Advertisement`) %>% 
  mutate(is_advertisement = if_else(is_advertisement == "YES", TRUE, FALSE)
       , quarter = quarter(date))
```

# Case 1

## Overview
```{r}
# quick summary statistics table
data1 %>% 
  describe()

```

### Volume Distributioin
```{r}
# create histogram of sales volume to show the distribution of sales
# across the dataset
weekly_sales_hist <- 
  data1 %>% 
  ggplot(mapping = aes(volume)) +
  # geom_density()
  geom_histogram(color = "white"
                , binwidth = 10000
                , boundary = 150000
                , fill = "#0067a6") +
  xlab("Volume (Units)") +
  ylab(element_blank()) +
  labs(title = "Weekly Sales Volume Distribution") 

# save plot as .png
ggsave("output/weekly_sales_hist.png"
     , plot = weekly_sales_hist
     , width = 7
     , height = 5
     )
```
### Volume Distribution as a Function of price
```{r}
# show distribution of sales volume for each price point
# the rain cloud plot shows the shape of the distribution 
# and also demonstrates that a t-test would likely 
# fail to reject the null hypothesis
vol_price_raincloud <- 
data1 %>% 
  mutate(fct_price = factor(price)) %>% 
  ggplot(mapping = aes(x = fct_price, y = volume)) +
  geom_half_boxplot(side = "l") +
  geom_half_point(side = "l", aes(color = fct_price)) +
  scale_color_manual(values = c("#0464a4", "#b7d4e6", "#5f9fc8", "#7cb4d4")) +
  geom_half_violin(side = "r", aes(fill = fct_price)) +
  scale_fill_manual(values = c("#0464a4", "#b7d4e6", "#5f9fc8", "#7cb4d4")) +
  scale_x_discrete(breaks = c("4.96", "4.98",
                              "4.99", "5")
                 , labels = c("$4.96", "$4.98",
                              "$4.99", "$5.00")) +
  xlab(label = "Price") +
  ylab(label = "Volume") +
  labs(title = "Volume Distribution by Price") +
  theme(legend.position = "none") +
  coord_flip()

# save plot as .png
ggsave("output/vol_price_raincloud.png"
     , plot = vol_price_raincloud
     , width = 7
     , height = 5
     )
  
```
### Hypothesis testing
```{r t-test-price-volume}
# t-test statistically showing what is visually shown in the previous chunk
# there are no statistical differences in sales volume
t.test(
    data1 %>% filter(price == 4.99) %>% select(volume)
  , data1 %>% filter(price != 4.99) %>% select(volume)
  , alternative = "two.sided"
  ) #%>%  tidy()
```
### Price Distribution
```{r price-histogram}
# histogram to show the count of records at each price point
price_dist <- 
data1 %>% 
  ggplot(mapping = aes(price)) +
  geom_histogram(bins = 5
               , color = "white"
               , fill = "#0067a6") +
  scale_x_continuous(labels = label_dollar()) +
  xlab("Price") +
  ylab(element_blank()) +
  labs(title = "Price Distribution") 

# save plot as .png
ggsave("output/price_dist.png"
     , plot = price_dist
     , width = 7
     , height = 5
     )
```
### Volume Over Time
```{r}
# plot sales volume over time, each color represents a different price point
# the line shows the relationship between sales volume and time
vol_time <- 
data1 %>% 
  ggplot(mapping = aes(x = date, y = volume)) +
  geom_smooth(method = "lm") + #"gam", formula = y ~ s(x, bs = "cs")) +
  geom_point(aes(color = factor(price))) +
  scale_color_manual(values = c("#ff6078", "#ae61b7", "#0067A6", "#ffa600")
                    , breaks = c("4.96", "4.98",
                                 "4.99", "5")
                    , labels = c("$4.96", "$4.98",
                                 "$4.99", "$5.00")
                    , name = "Price") +
  scale_x_datetime(date_breaks = "3 months", date_labels = "%m/%y") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  labs(title = "Sales Volume Over Time") +
  xlab("Date") +
  ylab("Volume (Units)") +
  theme(legend.position = "bottom")

# save plot as .png
ggsave("output/vol_time.png"
     , plot = vol_time
     , width = 7
     , height = 5
     )

```
## Modeling
```{r model}
# Standard Level OLS 
elasticity_model2 <- lm(volume ~ price + date, data = data1)
summary(elasticity_model2)

```

```{r volume-model}
# elasticity model
predicted_vol_model <- lm(log(volume) ~ log(price), data = data1)

# create table with predicted volume based on price.
# include Potential promotional price as well as price points between 
# lowest price in the data and promotional price
predicted_volume <- 
data1 %>% 
  group_by(price) %>% 
  summarise(volume = max(volume)) %>% 
  bind_rows(tibble(price = c(4.25, 4.5, 4.75), volume = NULL)) %>% 
  arrange(price) %>% 
  bind_cols(predict(predicted_vol_model, predicted_volume
                    , se.fit=TRUE, interval="confidence")$fit) %>% 
  mutate(pred_vol = exp(fit)
       , pred_lower = exp(lwr)
       , pred_upper = exp(upr)
       , pred_rev = pred_vol*price
       , not_in_model = if_else(is.na(volume), TRUE, FALSE))
  # mutate(log_price = log(price), log_volume = log(volume))
```

```{r pred-vol-plot}
# simple line plot showing the statistically insignificant "predicted" values
# plot actual data on top for context
pred_plot <- 
  ggplot(data = predicted_volume, 
         mapping = aes(x = pred_vol, y = price)) +
  geom_line(color = "red") +
  geom_point(data = data1, mapping = aes(x = volume, y = price)
             , color = "#0067A6", size = 0.5, alpha = 0.5) +
  scale_y_continuous(name = "Price", n.breaks = 10
                     , labels = scales::label_dollar()) +
  scale_x_continuous(name = "Volume (Units)"
                     , labels = label_number(scale_cut = cut_short_scale())
                     , n.breaks = 7) +
  labs(title = "Predicted Volume")

# save plot as .png
ggsave("output/pred_plot.png"
     , plot = pred_plot
     , width = 7
     , height = 5
     )
```

```{r pred-plot-conf}
# simple line plot showing the statistically insignificant "predicted" values
# plot actual data on top for context
# additional context added, confidence interval. this show that the promo price
# is too far away from the data to make a valid prediction.
pred_plot_conf <- 
ggplot(data = predicted_volume, 
         mapping = aes(x = pred_vol, y = price)) +
  geom_line(color = "red") +
  geom_ribbon(aes(xmin = pred_lower, xmax = pred_upper)
              , alpha = 0.25) +
  geom_point(data = data1, mapping = aes(x = volume, y = price)
             , color = "#0067A6", size = 0.5, alpha = 0.5) +
  scale_y_continuous(name = "Price", n.breaks = 10
                     , labels = scales::label_dollar()) +
  scale_x_continuous(name = "Volume (Units)"
                     , labels = label_number(scale_cut = cut_short_scale())
                     , n.breaks = 10) +
  labs(title = "Predicted Volume with Confidence Interval")

# save plot as .png
ggsave("output/pred_plot_conf.png"
     , plot = pred_plot_conf
     , width = 7
     , height = 5
     )
```



# Case 2

## Data Overview
```{r correlation-table}
# create cor table
data2 %>% 
  select(2:7) %>% 
  cor()
```

```{r advertisement-revenue-ttest}
t.test(
    data2 %>% filter(is_advertisement == TRUE) %>% select(revenue)
  , data2 %>% filter(is_advertisement == FALSE) %>% select(revenue)
  , alternative = "two.sided"
  )
```

```{r rev-shelf}
# plot the relationship between revenue and additional shelf space
# additionally, color points by advertisement. which is a significant predictior
rev_shelf <- 
data2 %>% 
  ggplot(mapping = aes(x = shelf_space_pct
                     , y = revenue
                     , color = is_advertisement)) +
  scale_color_manual(values = c("#0464a4", "#ff6078")
                   , name = "Any Store Advertisement"
                   , labels = c("No", "Yes")) + # "#5f9fc8" "#7cb4d4"
  geom_smooth(method = "lm") +
  geom_point() +
  scale_x_continuous(name = "Additional Shelf Space (%)"
                   , labels = label_percent(scale = 1)
                   , n.breaks = 5) +
  scale_y_continuous(name = "Revenue"
                   , labels = label_dollar(scale = 0.000001
                                         , suffix = "M")) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  labs(title = "Revenue And Additional Shelf Space") +
  theme(legend.position = "bottom")


# save plot as .png
ggsave("output/rev_shelf.png"
     , plot = rev_shelf
     , width = 7
     , height = 5
     )
```

```{r case-2-model}
# multivariate OLS regression
case_2_model <- lm(log(volume) ~ log(price) + num_stores + date
                           + shelf_space_pct + is_advertisement
                  , data = data2) 
# summary(case_2_model) # results table of regression

# create fitted table with predicted values 
# this is done in two stages. This is stage 1, setting up the table
case_2_pred <-
data2 %>% 
  group_by(price) %>% 
  summarise(
    num_stores = mean(num_stores)
  , date = max(date)
  , shelf_space_pct = mean(shelf_space_pct)) %>% 
  bind_rows(tibble(price = c(4.25, 4.5, 4.75), volume = NULL)) %>% 
  mutate(
     num_stores = if_else(is.na(num_stores), 1000.0000, num_stores)
   , date = if_else(is.na(date), ymd("2016-12-18"), ymd(date))
   , shelf_space_pct = if_else(is.na(shelf_space_pct)
                             , 23.33333, shelf_space_pct)
   , is_advertisement = TRUE) %>% 
  arrange(price)

# Stage 1 fitted table. adding predictions to the table.
case_2_pred <- case_2_pred %>% 
  bind_cols(predict(case_2_model, case_2_pred
                    , se.fit=TRUE, interval="confidence")$fit) %>% 
  mutate(pred_vol = exp(fit)
       , pred_lower = exp(lwr)
       , pred_upper = exp(upr)
       , pred_rev = pred_vol*price)

# showing predicted values at $4.25
# not significant, but providing for posterity 
case_2_pred %>% 
  filter(price == 4.25) %>% 
  select(1, 6:12) 
# predicted volume and revenue if price $4.25 *Not Statistically Significant
```

```{r case-2-model-plot}
# plotting the multivariate model. This is done by using the mean value for all
# the other covariates, and only changing price (aka: ceteris paribus) 
# interesting note, the sign for price flips to positive. Which is why there is
# and upward sloping demand curve (which makes no sense)
ggplot(case_2_pred, mapping = aes(x = pred_vol, y = price)) +
  geom_line() +
  geom_smooth(data = data2, mapping = aes(x = volume, y = price))

```

```{r case-2-model2, eval=FALSE}
# this model drops date and is_advertisement. this simplified
# model retains the negative coefficient on price, but it is still 
# statistically insignificant. Therefore, this model is not included in the
# presentation.
case_2_model2 <- lm(log(volume) ~ log(price) + num_stores + shelf_space_pct 
                  , data = data2)
summary(case_2_model2)
```





# Unused Chunks
These are chunks of code that ended up not being used in the final presentation.
They include early ideas/tests or broken code. The chunk option as been set to 
eval=FALSE to ensure they do not run if this file is knitted.
```{r predict-rev-model, eval=FALSE}
predicted_volume %>% 
  mutate(predicted_revenue = predicted_volume*log(price))

```
```{r pred-price-model, eval=FALSE}
log_price_model <- lm(log(price) ~ log(volume), data = data1)
summary(log_price_model)

data1 %>% 
  group_by(price) %>% 
  summarise(volume = max(volume)) %>% 
  bind_rows(tibble(price = 4.25, volume = NULL)) %>% 
  add_predictions(log_price_model) %>% 
  mutate(pred = if_else(is.na(pred),log(4.25), pred)) %>% 
  mutate(volume = if_else(is.na(volume)
                        , exp((pred - log_price_model$coefficients[1])
                          /log_price_model$coefficients[2])
                          , volume)
         ) %>% 
  ggplot(aes(x = volume, y = price)) +
  geom_line() +
  geom_point(data = data1, mapping =  aes(x = volume, y = price)) 
```
```{r pred-price-plot, eval=FALSE}
data1 %>% 
  ggplot(mapping = aes(x = volume, y = price)) +
  geom_point() +
  geom_smooth(method = lm) 


```
```{r case-2-plot, eval=FALSE}
data2 %>% 
  group_by(num_stores) %>% 
  summarise(avg_rev = mean(revenue)) %>% 
  mutate(num_stores = factor(num_stores)) %>% 
  ggplot(mapping = aes(x = num_stores
                     , y = avg_rev)) +
  geom_col() +
  coord_flip()
```
```{r price-over-time, eval=FALSE}
data1 %>% 
  ggplot(mapping = aes(x = date, y = price)) +
  geom_line() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_datetime(date_breaks = "3 months", date_labels = "%m/%y")
```
```{r num-stores-histogram, eval=FALSE}
data2 %>% 
  ggplot(mapping = aes(num_stores)) +
  geom_histogram(bins = 6, boundary = 1000, color = "white")
```
```{r case-2-model2, eval=FALSE}
case_2_model2 <- lm(volume ~ price + num_stores + shelf_space_pct 
                  + is_advertisement + factor(quarter)
                  , data = data2)
summary(case_2_model2)
```
```{r case-2-model3, eval=FALSE}
case_2_model3 <- lm(volume ~ price + num_stores + date
                           + shelf_space_pct + is_advertisement
                           + I(price^2)
                  , data = data2)
summary(case_2_model3)
```